---
title: "Budweiser Presentation"
output: 
  html_document: 
    keep_md: yes
---
## Functional App

https://nedeinlein.shinyapps.io/Beer_App/

##  Introduction
Good morning.  
Today we will showcase the answers to the following questions based on the provided data:  
* Where are breweries located in the United States?  
* Where did we land post data clean up?  
* What is the median ABV and IBU in each state?  
* Who has the most alcoholic and most bitter beer?  
* What does the ABV distribution look like in the United States?  
* Is there a relationship between ABV and IBU?  
We will also look into our ability to successfully identify an IPA.  
Finally, we will showcase promising test markets for a new IPA.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message= FALSE)
library(tidyverse)
library(maps)
library(mapproj)
library(usmap)
library(curl)
library(class)
library(e1071)
library(caret)
library(plotly)
library(fuzzyjoin)
```


## Where are breweries located in the United States?
The graphic below provides a heat map of the breweries in all 50 states and Washington DC.  

```{r Map Breweries in Each State}
#load brewery data
breweries <- read.csv(curl("https://raw.githubusercontent.com/tricleve/6306_CaseStudy1/master/Breweries.csv"))
#Trim space in front of abbreviation
breweries$abbr = str_trim(breweries$State)
#Rename Brewery id column to enable merge with beers.csv
colnames(breweries)[1] = "Brewery_id"
#Generate brewery count by state
brewery_count_by_state = count(breweries,abbr)
#Clean up column name to be more descriptive
colnames(brewery_count_by_state)[2] = "Breweries"
#Merge/select brewery count with statepop (usmap data set) to get fips for plot.
plotbrew = merge(statepop,brewery_count_by_state,by = "abbr") %>% select(fips, Breweries)
#Plot brewery count on the state map.  Includes Alaska, Hawaii and DC
plot_usmap(data=plotbrew, values="Breweries", color = "black") + scale_fill_gradient(name = " Brewery Count", low = "yellow", high = "red") + labs(title="Brewery count by state", subtitle="All 50 states and DC") + theme(legend.position = "right")
```  
    
The code above cleans the brewery data, merges the brewery data with the "statepop" data set from the usmap library and finally plots the brewery counts by state as a heat map using the plot_usmap function.  

## Merging beer data with brewery data
```{r Merge Beer Data}
beers <- read.csv(curl("https://raw.githubusercontent.com/tricleve/6306_CaseStudy1/master/Beers.csv"))
#Full join to ensure we bring in all data from both sets
comb_data = full_join(beers,breweries, by="Brewery_id")
tail(comb_data, 6)
head(comb_data, 6)
```
  
The code above reads the beers.csv from github and completes a full join of beers.csv and breweries.csv.

## Where did we land post data clean up?
1,007 observations were removed from the data due to missing value in ABV, IBU and Style.  Remaining data is adequate for analysis with 1,403 observations remaining.
```{r Address Combined Data Missing Values}
sum(is.na(comb_data$ABV))
sum(is.na(comb_data$IBU))
sum(is.na(comb_data$Style))
beerclean = comb_data %>% filter (!is.na(ABV) & !is.na(IBU) & !is.na(Style))
dim(comb_data)
dim(beerclean)
```
  
The code above identifies the missing values in three fields(ABV, IBU and Style) and then removes them from the combined data set.

## What is the median ABV and IBU in each state?
The graphs below are a bar chart of the median ABV and median IBU for each state. The states are ordered alphabetically. The y axis for ABV has been shifted to make it easier to see differences in ABV between the states.  
```{r Plot median ABV and IBU by state}
## The two plots below present the states alphabetically.
beerclean %>% group_by(abbr) %>% summarise(medianABV= median(ABV))%>% ggplot(aes(abbr,medianABV, fill=medianABV))+geom_bar(stat="identity", width=.5)+ coord_cartesian(ylim=c(.03,.07)) + scale_fill_gradient(low = "yellow", high= "red")+ labs(title="Median ABV by State", x="State", y="Median ABV") + theme(axis.text.x = element_text(angle=65, vjust=.3)) 
beerclean %>% group_by(abbr) %>% summarise(medianIBU= median(IBU)) %>% ggplot(aes(abbr,medianIBU, fill=medianIBU))+geom_bar(stat="identity", width=.5) + scale_fill_gradient(low = "yellow", high= "red")+ labs(title="Median IBU by State", x="State", y="Median IBU") + theme(axis.text.x = element_text(angle=65, vjust=.3)) 
```
  
The code above plots the median ABV and IBU by state alphabetically.  The view of the y axis was shifted from 0.0 to 0.03 to provide an easier way to see the differences between states.  The color scheme was chosen to be consistent with all graphs.

## Which state has the most alcoholic and most bitter beer?
Louisville, KY has the most alcoholic beer.
Astoria, OR has the most bitter beer.
```{r Show states with highest ABV and IBU}
beerclean[grep(max(beerclean$ABV),beerclean$ABV),] %>% select(City, abbr, Name.y, Name.x, ABV)
beerclean[grep(max(beerclean$IBU),beerclean$IBU),] %>% select(City, abbr, Name.y, Name.x, IBU)
```
  
The code above finds the row of the highest ABV/IBU and then selects pertinent columns for a clean display.

## What does the ABV distribution look like in the United States?
The distribution of ABV shows a normal distribution with a right skew.
``` {r Summary and distribution of ABV}
beerclean %>% ggplot(aes(ABV)) + geom_histogram(fill ='red', binwidth = .005) + labs(title="Distribution of ABV")
summarise(beerclean, max=max(ABV), min=min(ABV), mean = mean(ABV), median = median(ABV), sd=sd(ABV))
```
  
The code above plots a histogram of ABV with a specific binwidth.  A dataframe is also printed to show the max, min, mean, median and standard deviation.

## Is there a relationship between ABV and IBU?
There appears to be a positive linear relationship between Bitterness and Alcohol Content.  As ABV increases, IBU increases.
``` {r Scatterplot IBU vs ABV}
IPA <- beerclean %>% filter(grepl("IPA",Style))
ALE <- beerclean %>% filter(grepl("Ale", Style))
LITE <-beerclean %>% filter(grepl("Pilsner", Style) | grepl("Lager", Style))
IPAC <- IPA %>% mutate(Beerclass = "IPA")
ALEC <- ALE %>% mutate(Beerclass = "ALE")
LITEC <- LITE %>% mutate(Beerclass = "Pilsner/Lager")
beertrain <- rbind(IPAC,ALEC,LITEC)
beertrain %>% ggplot(aes(ABV, IBU)) + geom_jitter(color='red', width=.01) + labs(title="Scatterplot Bitterness vs Alcohol", x="Alcohol Content (ABV)", y="Bitterness (IBU)")
```
  
The code above performs the initial categorization of styles for our KNN model, but plots all points in a uniform color.  Jitter was chosen over point because it gave a better visual of the observations.

## Can we create a model to identify a beer as an IPA?
We were able to create a model to identfy an IPA with 74% accuracy.  The sensitivity of the model is 76% and the specificity is 88%.
``` {r KNN to classify IPAs, Ales, Pilsner/Lager}
beertrain %>% ggplot(aes(ABV, IBU, color=Beerclass)) + geom_jitter(width=.01) + labs(title="Scatterplot Depicting Beer Style", x="Alcohol Content (ABV)", y="Bitterness (IBU)")
#KNN train
set.seed(6)
splitPerc = .75
trainIndices = sample(1:dim(beertrain)[1],round(splitPerc * dim(beertrain)[1]))
train = beertrain[trainIndices,]
test = beertrain[-trainIndices,]
#finding most accurate K value
accs = data.frame(accuracy = numeric(40), k = numeric(40))
for(i in 1:40)
{
  classifications = knn(train[,c(3,4)],test[,c(3,4)],train$Beerclass, prob = TRUE, k = i)
  table(test$Beerclass,classifications)
  CM = confusionMatrix(table(test$Beerclass,classifications))
  accs$accuracy[i] = CM$overall[1]
  accs$k[i] = i
}
p = plot(accs$k,accs$accuracy, type = "l", xlab = "k")
#knn
classifications = knn(train[,c(3,4)], test[,c(3,4)], train$Beerclass, prob = TRUE, k = 3)
table(classifications,test$Beerclass)
confusionMatrix(table(classifications,test$Beerclass))
```
  
The code above replots the data set showing the the categories of style( IPA, Ale, Pilsner/Lager).  The code then separates a training and test set, finds the most accurate K value, performs the kNN and prints the confusion matrix.

## What markets are good test areas for a new IPA?
In looking at the per captia consumption of beer by state, we found the top 10 states. These states would be a good test market for a new IPA.  The graphic and table below show these states.
``` {r Knock your socks off}
#Source for per capita consumption: Kaplan, Jacob. Apparent Per Capita Alcohol Consumption: National, State, and Regional Trends 1977-2018. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2020-06-29. https://doi.org/10.3886/E105583V4
beerrate <- read.csv(curl("https://raw.githubusercontent.com/tricleve/6306_CaseStudy1/master/apparent_per_capita_alcohol_consumption_1977_2018.csv"))
#filter to look at 2018
BR <- beerrate %>% filter(year == 2018)
#Select State, year and beer
BRR <- BR %>% select(state, year, number_of_beers)
#Column Rename
BRR <-BRR %>% rename(State = state, "Beers_per_person_2018" = "number_of_beers")
#pull state data for mapping
state.abb
state.name
df <- data.frame(state.abb,state.name)
df <- df %>% rename(State = state.name)
#merging state data
df2 <- regex_left_join(df,BRR, by = "State",ignore_case = TRUE)
df3 <- df2[-c(49,5),]
df4 <- df3[,-c(3,4)]
df4 <- df4 %>% rename(state = state.abb, State_Name = State.x)
df4 <- df4[order(-df4$Beers_per_person_2018),]
df5 <- df4[-c(11:50),]
#Top 10 states for per capita beer consumption
df5
#Plots
plot_usmap(data=df5, values="Beers_per_person_2018", color = "black") + labs(title = "Top Ten Beer Drinking States", legend.position = "center", caption = "Consumption source: https://doi.org/10.3886/E105583V4") + theme(legend.position = "right") + scale_fill_gradient(name = "Beers/person 2018", low = "yellow", high= "red")
```
  
This code grabs per capita consumption by state for 2018, merges it with state data and prints the top 10 states using plot_usmap.  The top 10 states is printed out with the consumption information as well.

## Conclusion
In Summary:  
* High competition states are Colorado, California, Oregon, and Michigan  
* Average of all beers show ABV at 5.5% and IBU > 30  
* Highest ABV was London Balling out of KY  
* Highest Bitterness was Bitter Bitch Imperial IPA out of OR  
* ABV is a normal curve with a right skew centered around 5.5%.  
* Proved Positive correlation between ABV and IBU  
* KNN can accurately classify beers by type for new releases  
* New England region is prime market for new IPA beer with low competition from other breweries  

## Thank You